I don't know if projection is categorical. \\ Did \citealt{mandelkern-etal2020} discover that it is?

### 2023-11-16 writing/naturalness-paper/paper.tex
updated naturalness paper

\documentclass[11pt,fleqn]{article}
\usepackage[margin=1in,top=1in,bottom=1in]{geometry}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
%\usepackage[dvips]{graphics}
%\usepackage[table]{xcolor}
%\usepackage{amssymb}
\usepackage{float}
%\usepackage{subfig}
\usepackage{booktabs}
\usepackage{subcaption}

@@

\title{I don't know if projection is binary. \\ Did \citealt{mandelkern-etal2020} discover that it is?}

@@

Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{frege1892,ccmg90}). While projection is often characterized as binary (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and proposed that naturalness ratings in explicit ignorance contexts are more suitable to measure binary projection and, thereby, to distinguish presupposition triggers from nontriggers. This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} can distinguish factive predicates (presumed presupposition triggers) from nonfactive ones (nontriggers). While the results do not support \citepos{mandelkern-etal2020} hypothesis and are rather in line with results based on inference rating measures (e.g., \citealt{degen-tonhauser-language}), the results suggest that naturalness ratings in explicit ignorance contexts are sensitive to felicity conditions of anaphoric projective content.

@@

\noindent
{\bf Keywords:} Presuppositions, binary vs.\ gradient projection, inference ratings, naturalness ratings in explicit ignorance contexts.

@@

\noindent
For helpful comments on this project we thank Craige Roberts, Gregory Scontras, and Mandy Simons.

@@

Presuppositions are defined as content that typically projects out of entailment-canceling environments, like the scope of negation or a polar interrogative (e.g., \citealt{frege1892,ccmg90}).\footnote{Conventional implicatures also project out of these environments. They differ from presuppositions in that they contribute novel information. For discussion see, e.g., \citealt{ccmg90,potts05}.} Thus, a content like the content of the clausal complement (CC) of {\em know} in (\ref{first}a) is taken to be a presupposition because a speaker who utters (\ref{first}a) is typically taken to believe the CC, that Julian dances salsa, even though the clausal complement occurs in a polar interrogative. By contrast, the CC of {\em think} in (\ref{first}b) is not assumed to be a presupposition because it does not typically project. Accordingly, {\em know} is analyzed as a presupposition trigger, but {\em think} is not.

@@

To investigate whether a particular content is a presupposition, empirical research has long relied on inference rating measures. Such measures involve presenting participants with an utterance (like that in (\ref{first}a)) and contents (like the content that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content. Many theoretical works on projection assume that projection is a binary property of utterance content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition). There is, however, mounting empirical evidence (based on investigations that used inference rating measures) that projection is not a binary property of  content but rather a gradient one, with  content being more or less projective. \citealt{karttunen71b}, for instance, already suggested that the contents of the clausal complements (CCs) of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}). Specifically, he suggested that a speaker who utters (\ref{kart}a) believes that the addressee has not told the truth, whereas a speaker who utters (\ref{kart}b) ``is not sure about the truth of the complement'' (p.63).

\begin{exe}
\ex\label{kart}
\begin{xlist}
\ex Did you regret that you had not told the truth?
\ex Did you discover that you had not told the truth? \hfill (\citealt[63]{karttunen71b})
\end{xlist}
\end{exe}
More recently, experimental investigations have also observed projection variability between presuppositions (e.g., \citealt{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CCs of {\em know, discover,} and {\em reveal} exhibit projection variability: The CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}. These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a binary, categorical one''. Further support for this proposal comes from  investigations of both factive and nonfactive predicates in \citealt{degen-tonhauser-language}. In a series of experiments and analyses of existing datasets, this work not only found projection variability between the CCs of factive predicates, but also no empirical support for a binary distinction in the projection of the CCs of factive and nonfactive predicates. This is because the CCs of both factive and nonfactive predicates were projective to varying degrees compared to nonprojective main clause content. A sample illustration of these results is given in Fig.~\ref{fig:dt1a} from their Exp.~1a, which shows the mean projection ratings of 5 factive predicates (in \color{orange}orange\color{black}) and 15 nonfactive predicates (in \color{green}green\color{black}).

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{figures/Language-Figure2}
\caption{Mean certainty rating (measuring projection) of main clause content and the CCs of the 20 predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language} (adapted by collapsing the nonfactive predicates into a single category). Error bars indicate 95\% bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the individual participants' ratings.}\label{fig:dt1a}
\end{figure}

Recently,  \citealt[\S6.2]{mandelkern-etal2020} challenged the hypothesis that projection is gradient by arguing that inference rating measures are not suitable to detect binary projection. Such measures are not suitable, they argued, because they do not distinguish between, on the one hand, contents that project because they are presuppositions and, on the other hand, contents that project because they are ``a natural conclusion to draw for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497). Inference rating measures, they suggested, might invite participants to draw projection inferences even when not lexically required because ``the content in question is right there before the [participants'] eyes" (p.498). They concluded that ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).

\citealt[\S6.2]{mandelkern-etal2020} suggested that naturalness ratings of utterances in explicit ignorance contexts are better suited to distinguish presuppositions from nonpresuppositional projection inferences.\footnote{\citealt{mandelkern-etal2020} refer to the measure as an acceptability rating measure, but since participants were asked to rate the naturalness of utterances, we refer to it as a naturalness rating measure.} On this measure, participants are presented with utterances in which a speaker states that they are ignorant about a content, as in (\ref{mtrig}a), and then follows up with an utterance in which the content occurs embedded in an entailment-canceling environment, such as the antecedent of a conditional in (\ref{mtrig}c). \citealt{mandelkern-etal2020} assumed that if the content is a presupposition (such as the pre-state content of {\em stop} in (\ref{mtrig}c)), the utterance is rated as unnatural in the explicit ignorance context because presuppositions are contents that the speaker believes to be true. That is, in such constellations,  participants ``will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).  If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'' as in the explicit ignorance contexts (p.497).\footnote{Presuppositions won't be relinquished, modulo local accommodation} Both presuppositions and nonpresuppositions were expected to be judged as natural in the support contexts, which entail the relevant content, illustrated in (\ref{mtrig}b).

\begin{exe}
\ex\label{mtrig} \citealt[490f.]{mandelkern-etal2020}
\begin{xlist}
\ex Explicit ignorance context: \\ Mary always was involved in a lot of sports, but I don't know whether she ever did any yoga.
\ex Support context: \\ Mary always was involved in a lot of sports, and she used to do yoga, too.
\ex Sentence with presupposition / nonpresupposition: \\ If Mary \{has stopped / now frowns on\} doing yoga, then Matthew will interview her for his story.
\end{xlist}
\end{exe}
\citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, only provided limited empirical evidence for the hypothesis that projection is a binary property of content. This is primarily because the goal of this work (including Exp.~3) was not to investigate this hypothesis, but rather whether presupposition projection in conjunctions is symmetric. The evidence provided in this work is also limited, however, because only four presupposition triggers and nontriggers each were investigated: The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i). It is possible that the pre-state content contributed by this expression is projective: (i), for instance, appears to give rise to the projective content that Mary did not frown on doing yoga in the past.

\begin{exe}
\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
\end{exe}
} In the control items of Exp.~3, these eight expressions were realized in (single-clause) antecedents of conditionals, as shown in (\ref{mtrig}c). Naturalness ratings were elicited in explicit ignorance and support contexts.

\citealt{mandelkern-etal2020} found (as shown in their Fig.~3) that the naturalness ratings for the nontriggers in the control items were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale). The presupposition triggers, on the other hand, had lower mean naturalness ratings in the explicit ignorance context (mean around 3) than in the support context (mean just above 5). Mandelkern and his colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492). They proposed (p.497):

@@

However, an inspection of the naturalness ratings for the eight (non)triggers provided in Fig.~8 in their Appendix 3 raises doubt about whether the measure indeed provides evidence for a binary distinction between presuppositions and nonpresuppositions. First, considering the mean naturalness ratings in the explicit ignorance contexts, we find ratings for the presuppositions range from just above 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}). As the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}), it is not clear whether naturalness ratings in explicit ignorance contexts provide evidence for a binary distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}. Second, the effect of context (explicit ignorance vs.\ support) also does not appear to be consistent for presuppositions or nonpresuppositions. Whereas the difference between the mean naturalness ratings in the explicit ignorance and support contexts was quite large for the (presuppositional) pre-state of {\em continue}, the difference was much less pronounced for the (presuppositional) CC of {\em be aware}, where it was roughly comparable to that of the (nonpresuppositional) pre-state of {\em frown on}. In sum, it is not clear that comparisons of ratings in support and explicit ignorance contexts provides empirical support for a binary projection either.

There also appears to be disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts. Whereas Mandelkern and his colleagues assumed that all presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only hard (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural. For instance,  {\em stop} and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic}c) with `??' instead of `$\#$'.} This disagreement in the literature necessitates further investigation.

\begin{exe}
\ex\label{eic2}
\begin{xlist}
\ex I have no idea whether Jane ever smoked, but she hasn't stopped smoking. \hfill (\citealt[443]{simons01})
\ex I have no idea whether John ended up participating in the
Road Race yesterday. But if he won it, then he has more victories than anyone else in history. \hfill (\citealt[39]{abusch10})
\end{xlist}
\ex\label{eic3}
\begin{xlist}
\ex\infelic I don't know if Jane ever rented ``Manhattan'' before, but perhaps she's renting it again. \hfill (\citealt[443]{simons01})

@@

This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} provides evidence that projection is a binary property of content rather than a gradient one.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated. The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft, to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.

One change we made to the design of \citepos{mandelkern-etal2020} Exp.~3 was that our experiment implemented a three-level context condition instead of a two-level one. Specifically, we replaced \citepos{mandelkern-etal2020} support context (which entailed the relevant context, see (\ref{mtrig}b)) with two contexts that were compatible with the relevant content but did not entail it. The two contexts differed in the prior probability of the relevant content: In the `lower prior probability' context, illustrated in (\ref{prior}a), the  relevant content (in (\ref{prior}c), that Julian dances salsa) had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context, illustrated in (\ref{prior}b). These two contexts were normed in \citealt{degen-tonhauser-openmind}.

\begin{exe}
\ex\label{prior}
\begin{xlist}
\ex Julian is German. \hfill [lower prior probability]
\ex Julian is Cuban. \hfill [higher prior probability]
\ex Did Cole discover that Julian dances salsa?
\end{xlist}
\end{exe}
This change was implemented to allow for comparison to the result of \citealt{degen-tonhauser-openmind}, namely that the projection of a content is sensitive to the content's prior probability. Specifically, this work found that the higher the prior probability of a content, the stronger the projection inference. Under standard analyses of presuppositions, like \citealt{heim83} or \citealt{vds92}, presuppositions are by default globally accommodated in such contexts. Accordingly, such analyses predict that presuppositions are rated as natural in both lower and higher probability contexts, just as in \citepos{mandelkern-etal2020} support contexts.

%A challenge in the identification of presuppositions is that they only {\em typically} project: It has long been know that presuppositions need not project in any utterance in which the trigging expression is realized (e.g., \citealt{} REFERENCES). For example, in the naturally occurring example in (\ref{acc}) the presupposition of {\em know} (in the framed box) does not project but is (analyzed as) locally accommodated in the scope of the epistemic modal adverb {\em perhaps}. 
%
%
%\begin{exe}
%\ex\label{acc} Whatever the reason, she knows she can't just bounce back, and she decides to leave. For others, the decision is based purely on fear. Perhaps she \fbox{knows} that if she were to stay another day, her life or the lives of her children would be gone. \hfill (\citealt[85]{beaver-belly})
%\end{exe}
%
%
%In view of this potential for presuppositions to not project, investigations into the question of which utterance contents are presuppositions (and which ones are not) must

@@

To investigate whether the measure proposed in \citealt{mandelkern-etal2020} provides support for a binary distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive clause-embedding predicate), and rated the naturalness of the interrogative in the context of the declarative.

@@

We recruited 425 participants on Prolific. Due to a server error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-binary, 2 preferred to not disclose). The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%. The median time spent on the task was 6:24 minutes. Participants were paid \$1.78 (corresponding to an hourly pay of \$16.6).

@@

Participants were instructed to rate ``how natural the question sounds in the context of the statement''. As shown in Figure \ref{f:trials}, they were asked give their rating on a slider from `totally unnatural' (coded as 0) to `totally natural' (coded as 1). The experiment began with four practice trials to familiarize participants with the task (see Supplement \ref{a:fillerPractice} for details). After responding to the 30 trials, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

>>>>

\documentclass[11pt,fleqn]{article}
\usepackage[margin=1in,top=1in,bottom=1in]{geometry}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
%\usepackage[dvips]{graphics}
%\usepackage[table]{xcolor}
%\usepackage{amssymb}
\usepackage{float}
%\usepackage{subfig}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{colortbl}

@@

\newcommand*\rots{\multicolumn{1}{R{90}{.7em}}}% no optional argument here, please!

% positive coefficients/difference
\definecolor{yellow1}{RGB}{102,102,0} 
\definecolor{yellow2}{RGB}{153,153,0} 
\definecolor{yellow3}{RGB}{204,204,0} 
\definecolor{yellow4}{RGB}{255,255,204}
%\definecolor{green1}{RGB}{0, 158, 115} % >.95
%\definecolor{green2}{RGB}{55, 185, 141} % .85-.95
%\definecolor{green3}{RGB}{88, 214, 167} % .75-.85
%\definecolor{green4}{RGB}{119, 242, 194} % <.75

% negative coefficients/difference
\definecolor{purple1}{RGB}{150,57,109}
\definecolor{purple2}{RGB}{188,78,139}
\definecolor{purple3}{RGB}{215,149,186}
\definecolor{purple4}{RGB}{236,206,223}

\title{I don't know if projection is categorical. \\ Did \citealt{mandelkern-etal2020} discover that it is?}

@@

Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{ccmg90}). While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that naturalness ratings in explicit ignorance contexts are more suitable to measure categorical projection and, thereby, to distinguish presuppositions from nonpresuppositions. This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} can distinguish factive predicates (presumed presupposition triggers) from nonfactive ones (nontriggers). The results do not support \citepos{mandelkern-etal2020} claim as they do not provide evidence for a categorical distinction between presuppositions and nonpresuppositions. {\bf FIX: Furthermore, the results suggest that naturalness ratings in explicit ignorance contexts may be sensitive to a variety of semantic and pragmatic properties of utterances in addition to the felicity conditions of anaphoric projective content.}

@@

\noindent
{\bf Keywords:} Presuppositions, categorical vs.\ gradient projection, inference ratings, naturalness ratings in explicit ignorance contexts.

@@

\noindent
For helpful comments on this project we thank Craige Roberts, Gregory Scontras, Mandy Simons, and the audience at the syntax/semantics discussion group at the University of Stuttgart.

@@

Presuppositions are often characterized as content that typically projects out of entailment-canceling environments, like the scope of negation or a polar interrogative (e.g., \citealt{ccmg90}).\footnote{Conventional implicatures also project out of these environments. They differ from presuppositions in that they contribute novel information. For discussion see, e.g., \citealt{ccmg90,potts05}.} Thus, a content like the content of the clausal complement (CC) of {\em know} in (\ref{first}a) is taken to be a presupposition because a speaker who utters (\ref{first}a) is typically taken to believe the CC, that Julian dances salsa, even though the clausal complement occurs in a polar interrogative. By contrast, the CC of {\em think} in (\ref{first}b) is not assumed to be a presupposition because it does not typically project. Accordingly, {\em know} is analyzed as a presupposition trigger, but {\em think} is not.

@@

Many theoretical works assume that projection is a categorical property of utterance content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition). There is, however, mounting empirical evidence that projection is not a categorical property of  content but rather a gradient one, with content being more or less projective. The measure used to establish this evidence are inference rating measures: Such measures involve presenting participants with an utterance (like that in (\ref{first}a)) and contents (like the content that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content. \citealt{karttunen71b}, for instance, suggested that the CCs of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}). Specifically, he suggested that a speaker who utters (\ref{kart}a) believes that the addressee has not told the truth, whereas a speaker who utters (\ref{kart}b) ``is not sure about the truth of the complement'' (p.63).

\begin{exe}
\ex\label{kart}
\begin{xlist}
\ex Did you regret that you had not told the truth?
\ex Did you discover that you had not told the truth? \hfill (\citealt[63]{karttunen71b})
\end{xlist}
\end{exe}
More recently, experimental investigations have also observed projection variability between presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CCs of {\em know, discover,} and {\em reveal} exhibit projection variability: The CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}. These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a categorical, categorical one''. Further support for this proposal comes from  investigations of both factive and nonfactive predicates in \citealt{degen-tonhauser-language}. In a series of experiments and analyses of existing datasets, this work not only found projection variability between the CCs of factive predicates, but also no empirical support for a categorical distinction in the projection of the CCs of factive and nonfactive predicates. This is because the CCs of both factive and nonfactive predicates were projective to varying degrees compared to nonprojective main clause content. A sample illustration of these results is given in Fig.~\ref{fig:dt1a} from their Exp.~1a, which shows the mean projection ratings of 5 factive predicates (in \color{orange}orange\color{black}) and 15 nonfactive predicates (in \color{green}green\color{black}).

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/mean-certainty-by-predicateType}
\caption{Mean certainty rating (measuring projection) of main clause (`MC') content and the CCs of the 20 \color{orange}factive \color{black} and \color{green}nonfactive \color{black} predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language} (adapted by collapsing the nonfactive predicates into a single category). Error bars indicate 95\% bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the individual participants' ratings.}\label{fig:dt1a}
\end{figure}

Recently,  \citealt*[\S6.2]{mandelkern-etal2020} challenged the hypothesis that projection is gradient by arguing that inference rating measures are not suitable to detect categorical projection. Such measures are not suitable, they argued, because they do not distinguish between, on the one hand, contents that project because they are presuppositions and, on the other hand, contents that project because they are ``a natural conclusion to draw for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497). Inference rating measures, they suggested, might invite participants to draw projection inferences even when not lexically required because ``the content in question is right there before the [participants'] eyes" (p.498). They concluded that ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).

\citealt[\S6.2]{mandelkern-etal2020} suggested that naturalness ratings of utterances in explicit ignorance contexts are better suited to distinguish presuppositions from nonpresuppositional projection inferences.\footnote{\citealt{mandelkern-etal2020} refer to the measure as an acceptability rating measure, but since participants were asked to rate the naturalness of utterances, we refer to it as a naturalness rating measure.} On this measure, participants are presented with utterances in which a speaker states that they are ignorant about a content, as in (\ref{mtrig}a), and then follows up with an utterance in which the content occurs embedded in an entailment-canceling environment, such as the antecedent of a conditional in (\ref{mtrig}c). \citealt{mandelkern-etal2020} assumed that if the content is a presupposition (such as the pre-state content of {\em stop} in (\ref{mtrig}c)), the utterance is rated as unnatural in the explicit ignorance context because presuppositions are contents that the speaker believes to be true. That is, in such constellations,  participants ``will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).  If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'' as in the explicit ignorance contexts (p.497). Both presuppositions and nonpresuppositions were expected to be judged as natural in the support contexts, which entail the relevant content, illustrated in (\ref{mtrig}b).

\begin{exe}
\ex\label{mtrig} \citealt[490f.]{mandelkern-etal2020}
\begin{xlist}
\ex Explicit ignorance context: \\ Mary always was involved in a lot of sports, but I don't know whether she ever did any yoga.
\ex Support context: \\ Mary always was involved in a lot of sports, and she used to do yoga, too.
\ex Sentence with presupposition / nonpresupposition: \\ If Mary \{has stopped / now frowns on\} doing yoga, then Matthew will interview her for his story.
\end{xlist}
\end{exe}
\citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, only provided limited empirical evidence for the hypothesis that projection is a categorical property of content. This is primarily because the goal of this work (including Exp.~3) was not to investigate this hypothesis, but rather whether presupposition projection in conjunctions is symmetric. The evidence provided in this work is also limited, however, because only four presupposition triggers and four nontriggers were investigated: The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.
%\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i). It is possible that the pre-state content contributed by this expression is projective: (i), for instance, appears to give rise to the projective content that Mary did not frown on doing yoga in the past. 
%
%\begin{exe}
%\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
%\end{exe}
%} 
In the control items of Exp.~3, these eight expressions were realized in (single-clause) antecedents of conditionals, as shown in (\ref{mtrig}c). Naturalness ratings were elicited in explicit ignorance and support contexts.

\citealt{mandelkern-etal2020} found (as shown in their Fig.~3) that the naturalness ratings for the nontriggers in the control items were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale). The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean just above 5). Mandelkern and his colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492). They proposed (p.497):

@@

However, an inspection of the naturalness ratings for the eight (non)triggers provided in Fig.~8 in their Appendix 3 raises doubt about whether the measure indeed provides evidence for a categorical distinction between presuppositions and nonpresuppositions. First, considering the mean naturalness ratings in the explicit ignorance contexts, we find ratings for the presuppositions range from just above 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}). As the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}), it is not clear that naturalness ratings in explicit ignorance contexts provide evidence for a categorical distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}. Second, the effect of context (explicit ignorance vs.\ support) also does not appear to be consistent for presuppositions or nonpresuppositions. Whereas the difference between the mean naturalness ratings in the explicit ignorance and support contexts was quite large for the (presuppositional) pre-state of {\em continue}, the difference was much less pronounced for the (presuppositional) CC of {\em be aware}, where it was roughly comparable to that of the (nonpresuppositional) pre-state of {\em frown on}. Thus, it is not clear that comparisons of ratings in support and explicit ignorance contexts provides empirical support for a categorical projection either.

Adding to this, there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts. Whereas Mandelkern and his colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural. For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.} This disagreement about the naturalness of presuppositions in explicit ignorance contexts necessitates further investigation.

\begin{exe}
\ex\label{eic2}
\begin{xlist}
\ex I have no idea whether Jane ever smoked, but she hasn't stopped smoking. \hfill (\citealt[443]{simons01})
\ex Context: ``two people [\ldots] know that Henry is searching for Jane, but who don't themselves know where Jane is:'' \\ If Henry discovers that Jane is in New York, there'll be trouble. \hfill (\citealt[434]{simons01})
\ex I have no idea whether John ended up participating in the
Road Race yesterday. But if he won it, then he has more victories than anyone else in history. \hfill (\citealt[39]{abusch10})
\end{xlist}
\ex\label{eic3}
\begin{xlist}
\ex\infelic I don't know if Jane ever rented ``Manhattan'' before, but perhaps she's renting it again. \hfill (\citealt[443]{simons01})

@@

This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} provides evidence that projection is a categorical property of content rather than a gradient one.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated. The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.

One change made to the design of \citepos{mandelkern-etal2020} Exp.~3 was that the experiment implemented a three-level context condition instead of a two-level one.\footnote{Another change was that participants gave their naturalness ratings on a slider with endpoints labeled `totally unnatural' and `totally natural' rather than on a 7-point Likert scale with endpoints labeled `completely unnatural' and `completely natural'.}  Specifically, \citepos{mandelkern-etal2020} support context (which entailed the relevant context, see (\ref{mtrig}b)) was replaced with two contexts that were compatible with the relevant content but did not entail it. The two contexts differed in the prior probability of the  content: In the `lower prior probability' context, illustrated in (\ref{prior}a), the  relevant content (in (\ref{prior}c), that Julian dances salsa) had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context, illustrated in (\ref{prior}b). The two contexts for each of the 20 complement clauses that were combined with the 20 predicates were normed in \citealt{degen-tonhauser-openmind}.

\begin{exe}
\ex\label{prior}
\begin{xlist}
\ex Julian is German. \hfill [lower prior probability]
\ex Julian is Cuban. \hfill [higher prior probability]
\ex Did Cole discover that Julian dances salsa?
\end{xlist}
\end{exe}
This change in the context condition was implemented to allow for comparison to the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability. Specifically, this work found that the higher the prior probability of a content, the stronger the projection inference. Under standard analyses of presuppositions, like \citealt{heim83} or \citealt{vds92}, presuppositions are by default globally accommodated in contexts that are compatible with the presupposition, like the lower and higher prior probability contexts. Accordingly, such analyses predict that presuppositions are rated as natural in both lower and higher probability contexts, just as in \citepos{mandelkern-etal2020} support contexts.

@@

To investigate whether the measure proposed in \citealt{mandelkern-etal2020} provides support for a categorical distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive clause-embedding predicate), and rated the naturalness of the interrogative in the context of the declarative.

@@

We recruited 425 participants on Prolific. Due to a server error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-categorical, 2 preferred to not disclose). The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%. The median time spent on the task was 6:24 minutes. Participants were paid \$1.78, corresponding to an hourly pay of \$16.6.

@@

Participants were instructed to rate ``how natural the question sounds in the context of the statement''. As shown in Figure \ref{f:trials}, they were asked give their rating on a slider from `totally unnatural' (coded as 0) to `totally natural' (coded as 1). The experiment began with four practice trials to familiarize participants with the task (see Supplement \ref{a:fillerPractice} for details). After rating the 30 trials, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

\begin{figure}[h]
\centering
%\begin{subfigure}{0.49\textwidth}
\fbox{\includegraphics[width=.7\textwidth]{figures/trial-eic}}
%\caption{Trial with explicit ignorance context}
%\end{subfigure}
%\\[.1cm]
%\begin{subfigure}{0.49\textwidth}
%\fbox{\includegraphics[width=\textwidth]{figures/trial-low}}
%\caption{Trial with lower prior probability context}
%\end{subfigure} \hfill \begin{subfigure}{0.49\textwidth}
%\fbox{\includegraphics[width=\textwidth]{figures/trial-high}}
%\caption{Trial with higher prior probability context}
%\end{subfigure}
\caption{Sample trial with explicit ignorance context.}\label{f:trials}
\end{figure}
